How to select predictive models for decision-making or causal inference?

### 2023-12-17 how_to_select_causal_models.tex
iter

Prediction models have flourished because their predictions can be easily
demonstrated and validated on left-out data. But they require more
careful validation for decision making, using a metric accounting for the
putative intervention, the $R\text{-risk}$. Even when treated and
untreated population differ little, as in RCTs, the $R\text{-risk}$
brings a sizeable benefit. To facilitate better model selection, we provide Python
code\footnote{\url{https://github.com/soda-inria/causal_model_selection}}.
% Using the $R\text{-risk}$ does make evaluation more
% complicated not only because the procedure is more involved, but also
% because each intervention requires a dedicated evaluation. However, such
% off-policy evaluation remains much less costly than the recommended good
% practice of impact evaluation testing the ability of a prediction model
% to actually guide patient health \cite{hendriksen2013diagnostic}. Also,
This model-selection procedure puts no constraints on the models used to
build predictive models: it opens the door to evaluating a wide range of
models, from gradient boosting to convolutional neutral, or language
models.

>>>>

Increasingly complex prediction models integrating richer medical data
have flourished because their predictions can be easily
demonstrated and validated on left-out data. But using them to underpin a
decision on whether to treat or not requires more careful validation,
using a metric accounting for the
putative intervention, the $R\text{-risk}$. The $R\text{-risk}$
brings a sizeable benefit to select the most adequate model, even when
model development is based on treated and
untreated population with little differences, as in RCTs. To facilitate better model selection, we provide Python
code\footnote{\url{https://github.com/soda-inria/causal_model_selection}}.
% Using the $R\text{-risk}$ does make evaluation more
% complicated not only because the procedure is more involved, but also
% because each intervention requires a dedicated evaluation. However, such
% off-policy evaluation remains much less costly than the recommended good
% practice of impact evaluation testing the ability of a prediction model
% to actually guide patient health \cite{hendriksen2013diagnostic}. Also,
This model-selection procedure puts no constraints on the models used to
build predictive models: it opens the door to evaluating a wide range of
models, from gradient boosting to convolutional neutral, or language
models.
